---
title: "Practical Machine Learning Project"
author: "Long Pei"
date: "October 25, 2014"
output: html_document
---

##Load and process data

### 1. Load required packages and the original data.

```{r}
library(caret)
library(randomForest)
training_raw<-read.csv("pml-training.csv",na.strings= c("NA",""," "))
```
### 2. Get the clean data by removing columns containing NAs
```{r}
# clean the data to exclude columns with NAs 
count_NAs <- apply(training_raw, 2, function(x) {sum(is.na(x))})
training_clean <- training_raw[,which(count_NAs == 0)]
```
### 3. Trim the data to exclude identifier columns
```{r}
# remove identifier columns (first 7 columns)
training_data <- training_clean[8:length(training_clean)]
```
### 4. Seperate the training set into two parts (training and validate)
```{r}
# create a seperation of the data into training and validation
set.seed(123)
inTrain <- createDataPartition(training_data$classe, p=0.7, list=FALSE)
training<-training_data[inTrain,]
validate<-training_data[-inTrain,]
```

##Variable selection
###We first include all the columns in the tidy data in the training dataset and via randomForest method to get a rough model, then based on varImp function to rank the importance of all these factors and then select only top ten factors to apply an improved randomForest method fit model to the validation dataset.

```{r}
# check all the columns we now include and get a rough model with randomForest
dim(training)
rough_fit<-randomForest(classe ~. , data=training)
# a genral view of the rough fit
rough_fit
# use the varImp to rank the importancces of all the factors and refine our model
varImp(rough_fit)
# use rough_fit on the validate part to check performances
predictCrossVal <- predict(rough_fit, validate)
confusionMatrix(validate$classe, predictCrossVal)
```

##Build up the model
###Here we build up the improved mode fit to replace the rough fit on the validate part to check performances.

```{r}
# use top 10 variables to build up the model
refined_index<-order(varImp(rough_fit)$Overall,decreasing = TRUE)[1:10]
# reshape the 10 improved columns (plus 'classe')
improve_training<-training[,c(refined_index,53)]
model_fit<-randomForest(classe ~. , data=improve_training)
# a genral view of the rough fit
model_fit
# use model_fit on the validate part to check performances
predictCrossVal <- predict(rough_fit, validate)
confusionMatrix(validate$classe, predictCrossVal)
```

##Prediction on test data
```{r}
test_raw<-read.csv("pml-testing.csv",na.strings= c("NA",""," "))
count_NAs <- apply(test_raw, 2, function(x) {sum(is.na(x))})
test_clean <- test_raw[,which(count_NAs == 0)]
test_data <- test_clean[8:length(test_clean)]
test<-test_data[,c(refined_index)]
# make prodiction
answers <- predict(model_fit, test)
answers<-as.character(answers)

# pml_write_files function
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
